Text as Policy: \\ Measuring Policy Similarity through Bill Text Reuse

### 2016-09-22 manuscript/main.tex
updated paper for nosplit

\subsection{Detecting Text Reuse}
To assess the maximal similarity of policies proposed in two pieces of legislation, we seek to identify large segments of equivalent or highly similar text (i.e., text reuse between bills). Alignment algorithms for discovering similar long sequences of text stand in contrast to bag-of-words methods, which are based on the comparison of word occurrence frequencies in documents. Bag-of-words methods may be effective for capturing broad topical areas in legislation, but will provide results that are too coarse for precisely assessing policy overlap. For example, in modeling bills in the US Congress using statistical topic models, arguably the canonical bag-of-words method for text analysis, \citet{gerrish2011predicting} find that topic models for legislative text perform most effectively when parameterized with 64 topics to model a corpus of 4,447 bills. While 64 topics provides a fairly detailed categorization of the domains covered in legislation, it is unrealistic to think that this method can isolate individual policy proposals.  Considering \citeapos{gerrish2011predicting} results again, that would mean that the US Congress considers nearly seventy bills, on average, for each specific policy proposed.

@@

\subsection{Data}
We apply the SW algorithm to a database of US state bills, collected by
\citet{burgess2016legislative} and the Sunlight
Foundation\footnote{\url{http://sunlightfoundation.com/}}. This data base
contains approximately 500,000 bills from 2008 to 2015. Not all bills from all
states are available for the whole time period. Figure \ref{fig:bill_desc}
displays the year ranges and number of bills in the bill data base. This
collection of bills is based on all bills that are available through the
Sunlight Foundation's \url{openstates.org} API. \url{openstates.org} is a website maintained by the Sunlight Foundation, in order to increase transparency in state politics. The Sunlight Foundation uses web scrapers to access all bills that are available on the websites of legislatures in all US states. This includes enacted legislation as well as bills that are still in the legislative process, or where not enacted. %The database contains the bill text as well as additional information on the legislative process for the bill. We have the first and the last version of the bill text, the sponsor of the bill and all legislative action that was taken on the bill. %The available meta-data for the bills includes a time-stamp for introduction and approval of the bill, the name and party affiliation of the sponsor(s), the state and the bill id.

@@

From this procedure we obtain ca. $10^9$ individual alignments (that is, shared
text between two sections of two bills) which corresponds to about $2 * 10^6$ bill dyads
($1.3 * 10^6$ bill dyads excluding within-state alignments). Figure
\ref{fig:alignment_score_distribution} displays the cumulative frequency
distribution of the alignments. The vast majority of alignments are very small
(less than 10). We use the alignment score as the quantitative measure of policy similarity. However, as we discuss below, due to the presence of extreme outliers, we work with this measure on the log scale. In order to aggregate the alignments to the bill level, we sum the alignment scores for each section pair of the corresponding bill dyad. Table \ref{tab:alignment_examples} displays three examples of section alignments. The `Left Text' and `Right Text' columns display the optimal alignment for the two sections the alignment score is displayed in the `Score' column. This example shows the reuse of legislative text in three states (New Jersey, North Carolina and Georgia).

@@

In the second approach we describe, which we have implemented in one analysis below, we do not exclude boiler plate according to a classification dichotomy. Rather, we down-weight common alignments in calculating similarity between bills. Alignments are identified at the section level, meaning that we identify up to $n_1 \times n_2$ alignments between two bills, with $n_1$ and $n_2$ being the number of sections in the two bills. The similarity score between two bills is calculated as the natural log of the sum of the numeric score associated with each section-pair alignment. To adjust for common alignments, we divide the raw alignment score from the SW algorithm by the frequency with which that exact same alignment occurs across the entire database of section-pair alignment. In this approach we follow the common intuition in natural language processing that the informativeness of text is inversely related to the frequency with which that text occurs \citep{robertson2004understanding} (see, e.g., the regular application of term frequency–inverse document frequency (tf-idf) to pre- or post-process text-based datasets and results). This measure is different and potentially preferable in two ways. First, it does not require the use of human coders. Second, it places text on a continuum of uniqueness rather than dichotomizes it as boiler plate or not.

@@

For this analysis we additionally employed the boilerplate adjustment strategy
described above\footnote{We plan to use boiler plate adjustment for all
    analyses. We are experimenting with the boiler plate removal on this
    analysis because the smaller size of the data. The final evaluation will be
    conducted on a held out data set of ncsl tables that is not included in the
current sample}. Before aggregating the alignments to the bill-dyad level, their
score is weighted by the inverse frequency of the alignment in the set of all
alignments for this analysis\footnote{The exact procedure is to compile the
matching elements of the two aligned sequences to a string that can then be
matched. This means that mismatches and gaps are ignored, since they would make
the alignment text unique.}. This results in down-weighting of alignments that occur too often to
be considered relevant policy language. Table
\ref{tab:expl_align_freq} displays some example alignments with very high and
very low frequency. From analysis of these frequencies we find that all alignments
that occur very frequent are indeed boilerplate. However, there are also still
many boilerplate alignments that occur infrequently. The currently employed
approach therefore improves the signal in the alignment score, but could be
further improved by matching and aggregating not only equal but also similar alignments (e.g. using another alignment
step or bag-of-words similarity measures).

\begin{table}[]
\centering
\caption{Example alignment text of comparable length occurring with high and low frequency.}
\label{tab:expl_align_freq}
\begin{tabular}{lll}
\hline
Frequency & Matched Alignment Text                           & Rank   \\
\hline
1         & to the treasurer of the county in which the      & 1      \\
1         & shall be for purposes of this                    & 1      \\
1         & of the higher education act of 1965              & 1      \\
...       & ...                                              & ...    \\
7089      & be it enacted by the legislature of the state of & 244747 \\
19349     & be it enacted by the of the state of             & 244840 \\
21100     & is amended to read as follows                    & 244847 \\
\hline
\end{tabular}
\end{table}

Figure \ref{fig:ncsl_prec_rec} displays the performance of the alignment score
in predicting if two bills are in the same NCSL table. We present
precision---the proportion of bill pairs with a given alignment score that
exceed the threshold on the $x$-axis that are in the same NCSL table, and
recall---the proportion of same-NCSL-table pairs that exceed a given alignment
score threshold. We display the results for the alignment scores adjusted for
boilerplate (left panel) and the unweighted scores (right panel). We see
moderate to high precision --- 60\% - 100\% with moderate
(higher than 6) to high alignment scores (about 1\% or 2000 of bill-dyad alignment
scores fall into this category) for the adjusted scores. The
threshold-classification using the unweighted scores performs much worse. This
is due to the fact that some bills with large numbers of section have high
bill-dyad level alignment scores simply due to the fact that they have many
boilerplate alignments. These results indicate that down-weighting by inverse
frequency is a promising direction to take.

Recall is low at thresholds that produce reasonable precision (above 6), but that is expected for our case, as
(1) the introduction of similar policies in bills in tow different states is a
rare event, and relatedly, (2) the appearance of two bills in the same NCSL
table is a relatively rare event (8\% in the current analysis); and classifier recall is notoriously low in rare event data \citep{weiss2000learning}---even commonly approaching zero, as in our case \citep{weiss2004mining}. Our results suggest that the presence of a large alignment score between bills is an effective indicator of policy similarity.

\begin{figure}
\centering     %%% not \center
\subfigure[Precision recall curve for boilerplate adjusted
scores]{\label{fig:prec_rec_adj}\includegraphics[width=0.4\textwidth]{figures/ncsl_prec_rec_adj.png}}
\subfigure[Precision recall curve for unweighted
scores]{\label{fig:prec_rec_old}\includegraphics[width=0.4\textwidth]{figures/ncsl_prec_rec_old.png}}
\caption{Precision-Recall curve for classifying two bills being in the same
    ncsl table by thresholding the bill-dyad alignment score. The left panel displays the curve for the alignment scores weighted by the inverse frequency of the alignment, the left panel displays the curve for the unweighted scores. The x axis
displays the threshold, the y axis represents precision and recall. The shaded
area in the background displays the frequency distribution of the alignment
scores.}
\label{fig:ncsl_prec_rec}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Identity Link} & \multicolumn{2}{c}{Log Link} \\
  \hline
 & Coefficient & p-value & Coefficient & p-value \\ 
  \hline
Intercept & -6760712.2229 & 0.0048 & 8.2122 & 0.0000 \\ 
  Diffusion Tie & 498466.2990 & 0.0496 & 0.2683 & 0.0968 \\ 
  Coverage & 2199070.8445 & 0.0000 & 1.4324 & 0.0000 \\ 
   \hline
\end{tabular}
\caption{Predicting number of alignments in legislation across states with diffusion ties. Coefficients calculated with OLS regression. $p$-values based on 5,000 QAP permuations.}
\label{tab:qap.diffusion}
\end{table}

@@

In the following sections we present several analyses to assess this correlation. When we are generating the alignments the bills are split up into sections. We generate an alignment score for a bill dyad by first taking the logarithm and then summing over all section dyads of the bill dyad. The ideal point of the bill is derived from the ideal point of the primary bill sponsor. For bills with several primary sponsors we use the average of those sponsors' ideal points to obtain a single measure. Figure \ref{fig:ideology_plot} displays the distribution of ideological distance and log-alignment score for the roughly 50 million dyads in the dataset. Each dot is a hexagonal tile, the color of the tile represents the number of dyads in that area. Two major observations can be made in this plot. First, the vast majority of alignment scores are close to zero. Second, the triangular shape of the distribution indicates the high precision - low recall character of the alignment score discussed above: few bills exhibit meaningful policy overlap with other bills, but in the region of lower ideological distance we observe a greater occurrence of positive outliers, which arise from significant policy overlap.

In order to investigate this pattern more closely, we additionally calculate
quantile regressions for the relationship between the alignment score and the
ideological distance. Figure \ref{fig:quantile_regression} displays the
coefficients from quantile regressions on a sequence of quantiles from the
median to the 0.995th quantile. Confidence intervals are calculated by resampling bills, which represents an implementation of the clustered nonparametric bootstrap proposed by \citet{harden2011bootstrap}. We see that there is no relationship between the ideological similarity of bill sponsors and the median alignment scores between bills. However, as the quantile we model increases---especially above the 90th percentile, we see that the relationship grows strong and negative in magnitude, and is statistically significant. These results both validate the use of text reuse as a measure of policy similarity, and reinforce the high precision and low recall nature of text reuse as a measure of policy similarity.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/quantile_regression.png}
    \caption{Coefficients of quantile regression for alignment score on
    ideological distance. Dots indicate the point estimate. Horizontal bar
represent 95\% confidence intervals obtained via the clustered bootstrap, [using
500 iterations]. Note that the x-axis is not in a natural scale}
    \label{fig:quantile_regression}
\end{figure}

>>>>

\subsection{Detecting Text Reuse}
\label{sec:d_t_r}
To assess the maximal similarity of policies proposed in two pieces of legislation, we seek to identify large segments of equivalent or highly similar text (i.e., text reuse between bills). Alignment algorithms for discovering similar long sequences of text stand in contrast to bag-of-words methods, which are based on the comparison of word occurrence frequencies in documents. Bag-of-words methods may be effective for capturing broad topical areas in legislation, but will provide results that are too coarse for precisely assessing policy overlap. For example, in modeling bills in the US Congress using statistical topic models, arguably the canonical bag-of-words method for text analysis, \citet{gerrish2011predicting} find that topic models for legislative text perform most effectively when parameterized with 64 topics to model a corpus of 4,447 bills. While 64 topics provides a fairly detailed categorization of the domains covered in legislation, it is unrealistic to think that this method can isolate individual policy proposals.  Considering \citeapos{gerrish2011predicting} results again, that would mean that the US Congress considers nearly seventy bills, on average, for each specific policy proposed.

@@

\subsection{Data}
We apply the SW algorithm to a database of US state bills, collected by
\citet{burgess2016legislative} and the Sunlight
Foundation\footnote{\url{http://sunlightfoundation.com/}}. This data base
contains approximately 500,000 bills from 2008 to 2015\footnote{Since
    computation is very demanding for this problem, the results reported in this
    version of the paper are based on a random sample of about half of the
    bills. We are in the process of completing the analyses for the whole
dataset.}. Not all bills from all
states are available for the whole time period. Figure \ref{fig:bill_desc}
displays the year ranges and number of bills in the bill data base. This
collection of bills is based on all bills that are available through the
Sunlight Foundation's \url{openstates.org} API. \url{openstates.org} is a website maintained by the Sunlight Foundation, in order to increase transparency in state politics. The Sunlight Foundation uses web scrapers to access all bills that are available on the websites of legislatures in all US states. This includes enacted legislation as well as bills that are still in the legislative process, or where not enacted. %The database contains the bill text as well as additional information on the legislative process for the bill. We have the first and the last version of the bill text, the sponsor of the bill and all legislative action that was taken on the bill. %The available meta-data for the bills includes a time-stamp for introduction and approval of the bill, the name and party affiliation of the sponsor(s), the state and the bill id.

@@

From this procedure we obtain ca. $1.2 * 10^8$ individual alignments (that is, shared
text between two bills). Figure \ref{fig:alignment_score_distribution} displays the cumulative frequency
distribution of the alignments. The vast majority of alignments are very small
(less than 50). Table \ref{tab:alignment_examples} displays three examples of alignments. The `Left Text' and `Right Text' columns display the optimal alignment for the two sections the alignment score is displayed in the `Score' column. This example shows the reuse of legislative text in three states (New Jersey, North Carolina and Georgia).

@@

In the second approach we describe, which we have implemented in one analysis
below, we do not exclude boiler plate according to a classification dichotomy.
Rather, we down-weight alignments alignments that consist of language that is
very common to all alignments. The underlying logic is that language that
appears in a large number of other alignments is probably not related to
specific policies. For examples alignments consisting of many stop words such as
`the' or `a' or of common procedural language like `legislature` or `section'
are most likely part of phrases that are either very frequent in common language
or very frequent in the legislative context. 
In order to quantify the similarity to common alignments, we calculate the
cosine similarity of the bag-of-words vector of each alignment with a random
sample of 1000 alignments\footnote{We rely on a random sample for computational
reasons, calculating the complete similarity matrix is computationally
unfeasible and unnecessary.}. The adjusted score for two bills $Y$ and $Y$ is
then:

\begin{equation}
    S^*_{X,Y} = S_{X,Y} \left(1 -
    \frac{1}{1000}\sum_{i=1}^{1000}\frac{\mathbf{\mathcal{A}_{X,Y}} \bullet
    {\mathbf{\mathcal{A}_i}}}{\lVert\mathbf{\mathcal{A}_{X,Y}\rVert\lVert\mathbf{\mathcal{A}_i}}\rVert}\right)
\end{equation}

Where $S_{X,Y}$ is the alignment score as described in Section \ref{sec:d_t_r},
$\mathbf{\mathcal{A}_j}$ is the bag-of-word vector representation of the
matching alignment text and $i = 1,\dots,1000$ are the sample alignments.

In this approach we follow the common intuition in natural language processing that the informativeness of text is inversely related to the frequency with which that text occurs \citep{robertson2004understanding} (see, e.g., the regular application of term frequency–inverse document frequency (tf-idf) to pre- or post-process text-based datasets and results). This measure is different and potentially preferable in two ways. First, it does not require the use of human coders. Second, it places text on a continuum of uniqueness rather than dichotomizes it as boiler plate or not.

@@

Figure \ref{fig:ncsl_prec_rec} displays the performance of the alignment score
in predicting if two bills are in the same NCSL table. We present
precision---the proportion of bill pairs with a given alignment score that
exceed the threshold on the $x$-axis that are in the same NCSL table, and
recall---the proportion of same-NCSL-table pairs that exceed a given alignment
score threshold. We see very high precision when setting the threshold to
relatively high scores ($>100$). Recall is low at thresholds that produce
reasonable precision (above 100), but that is expected for our case, as
(1) the introduction of similar policies in bills in tow different states is a
rare event, and relatedly, (2) the appearance of two bills in the same NCSL
table is a relatively rare event (8\% in the current analysis); and classifier
recall is notoriously low in rare event data \citep{weiss2000learning}---even
commonly approaching zero, as in our case \citep{weiss2004mining}. Our results
suggest that the presence of a large alignment score between bills is an
effective indicator of policy similarity. Futher more, low recall has

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/ncsl_pr_nosplit.png}
\caption{Precision-Recall curve for classifying two bills being in the same
    ncsl table by thresholding the bill-dyad alignment score. The x axis
displays the threshold, the y axis represents precision and recall. The shaded
area in the background displays the frequency distribution of the alignment
scores.}
\label{fig:ncsl_prec_rec}
\end{figure}

@@

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Wed Sep 21 20:12:14 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{Identity Link} & \multicolumn{2}{c}{Log Link} \\
  \hline
 & Coefficient & p-value & Coefficient & p-value \\ 
  \hline
Intercept & -33577305.97 & 0.00 & 10.04 & 0.00 \\ 
  Diffusion Tie & 2423709.31 & 0.05 & 0.27 & 0.10 \\ 
  Coverage & 10938819.90 & 0.00 & 1.38 & 0.00 \\ 
  \hline
\end{tabular}
\caption{Predicting number of alignments in legislation across states with diffusion ties. Coefficients calculated with OLS regression. $p$-values based on 5,000 QAP permuations.}
\label{tab:qap.diffusion}
\end{table}

@@

In the following sections we present several analyses to assess this correlation. The ideal point of the bill is derived from the ideal point of the primary bill sponsor. For bills with several primary sponsors we use the average of those sponsors' ideal points to obtain a single measure. Figure \ref{fig:ideology_plot} displays the distribution of ideological distance and log-alignment score for the roughly 50 million dyads in the dataset. Each dot is a hexagonal tile, the color of the tile represents the number of dyads in that area. Two major observations can be made in this plot. First, the vast majority of alignment scores are close to zero. Second, the triangular shape of the distribution indicates the high precision - low recall character of the alignment score discussed above: few bills exhibit meaningful policy overlap with other bills, but in the region of lower ideological distance we observe a greater occurrence of positive outliers, which arise from significant policy overlap.

In order to investigate this pattern more closely, we additionally calculate
quantile regressions for the relationship between the alignment score and the
ideological distance. Figure \ref{fig:quantile_regression} displays the
coefficients from quantile regressions on a sequence of quantiles from the
median to the 0.995th quantile. Confidence intervals are calculated by resampling bills, which represents an implementation of the clustered nonparametric bootstrap \citet[see e.g.][]{harden2011bootstrap}. Coefficients significant at the 95\% level are displayed in black, insignificant coefficients are grey. We see that there is only a very weak relationship between the ideological similarity of bill sponsors and the median alignment scores between bills. However, as the quantile we model increases we see that the relationship grows strong and negative in magnitude, and is statistically significant. These results both validate the use of text reuse as a measure of policy similarity, and reinforce the high precision and low recall nature of text reuse as a measure of policy similarity.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/quantile_regression.png}
    \caption{Coefficients of quantile regression for alignment score on
    ideological distance. Dots indicate the point estimate. Horizontal bar
represent 95\% confidence intervals obtained via the clustered bootstrap, [using
1200 iterations]. Note that the x-axis is not in a natural scale}
    \label{fig:quantile_regression}
\end{figure}
